### 2023-02-09 article.tex
Progress

\setcounter{chapter}{7}
\chapter{Research code is paradata and as vital as the article itself}

\abstract{

Paradata is data about the data collection process,
that allows use and reuse of data.
Within the context of computational research,
computer code is the paradata of an experiment,
allowing the study to be reproduced.
A recent study recommended how to make paradata (more) useful,
for paradata in general.
This study applies those recommendations to computer code,
using the field of genetic epidemiology
as an example.
The chapter concludes by some rules how to better code to serve as paradata,
and hence allowing computational research to be truly reproducible.

} % abstract

{\bf Keywords:} paradata, reproducible research, code, software,
FAIR data, computational research, Open Science, best practices,
genetic epidemiology

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Definitions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[h]
  \begin{tabular}{lp{5cm}p{5cm}}
    Term      & Definition                                            & Example                                                         \\
    \hline
    Code      & Body of text to be run by a computer                  & The scripts run by a computational experiment                   \\
    Data      & Individual facts, statistics, or items of information & A SNP that has a significant association                        \\
    Genotype  & The DNA allele at a certain location                  & AA, AC, CG, GT, ...                                             \\
    Paradata  & Data that describe the process of generating data     & The code to conclude that a SNP has a significant association   \\
    Phenotype & How an organism looks like in the broadest sense      & The concentration of IL6RA in the blood                         \\
    Primary data & Data that are worked with in a research            & The genotype of individuals   \\
    Metadata  & Data that provide information about other data        & The article that describes an experiment                        \\
    Raw data  & Data directly from the real world, that are not yet ready to be primary data & The raw results of getting the genotypes of individuals   \\
    Trait     & A phenotype                                           & The concentration of IL6RA in the blood                         
  \end{tabular}
  \caption{Terms used in this paper, and their definitions}
  \label{tab:definitions}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\epigraph{
  Talk is cheap. Show me the code.
}{
  Linus Torvalds, 2000-08-25
}

\paragraph{Simple example}

% Alex and Blake?
Two different researchers in genetic epideomology (more on that field later), 
write two equally good manuscripts
that describe an experiment with computational steps.
Both manuscripts are accepted by an equally prestigous journal after peer view. 
One researcher, however, does not supply the
computer code (from now on: 'code') that was used to generate the results,
where the other does.
Are the conclusion of these papers to be trusted equally?
Is this difference relevant and worth the effort?
This chapter discusses why code, a type of paradata, 
should be supplied
and how it should be supplied,
and claims that code is just as important as the paper itself.

\paragraph{The first use of paradata}

The concept of paradata (although not named as such yet) 
was introduced by \cite{couper1998measuring},
who developed a computer-assisted interview program
that, among others, records all key strokes,
measures the time used to answer each question 
and even the time the monitor is turned off.
The goal in that context was to assess and improve survey quality.

\paragraph{Definition of paradata}

This paper defines paradata as 'data about the data collection 
process' \cite{choumert2019using}.
This definition is chosen for its simplicity, 
yet there are multiple related and/or more nuanced 
definitions of paradata (see for example \cite{huvila2022improving} 
and \cite{skold2022interrogating}).

\paragraph{Code is paradata}

The code used in computational experiments is just that, 
'data about the data collection 
process', as it commonly downloads data, selects relevant subsets in those
downloaded data and performs statistical tests to generate results.
Code, hence, is paradata and this chapter explores and illustrates the
consequence of that premise.

\paragraph{Making paradata useful}

A recent paper explores the use of paradata to increase the impact of data,
stating that lack of paradata can be seen as 'a drastic constraint'
in the use of data and offer some suggestions to 
make paradata useful for data re(use):
paradata should be comprehensive, documented in a useful way, 
the documention and data should have co-evolved, 
and the paradata should be good to computers \cite{huvila2022improving}.
These suggestions appear to work well for code too
and will be discussed in detail below.

\paragraph{The goal of good paradata}

There are multiple reasons why useful paradata matter. 
Most obvious is that having useful paradata gives an understanding how
data is produced.
This insight helps researchers from different fields to understand each other
and collaborate.
Additionally, useful open data is needed for Open Science to convincingly show
its benefits.
Finally, in computational fields, 
it can help understand how scholarly knowledge is produced
\cite{huvila2022improving}.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{figure_1.png}
  \caption{
    Left: general relation between data, paradata and metadata.
    Right: the same relations specified for genetic epidemiology.
  }
  \label{fig:figure_1}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Genetic epidemiology}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\paragraph{Introduction}

This paper uses genetic epidemiology as a specific example,
to illustrate that the code is paradata,
and that this type of paradata is relevant.
However, any field that uses computation 
on sensitive data in its experiments,
could be used as an example.

\paragraph{How commonly code is published}\label{para:how-commonly-code-is-published}

Going back to the example scientists in the introduction,
where one does and the other does not publish the code,
it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
192 of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analysed the reproducibility of registered reports
in the field of psychology, 
where 37 out of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert, 
yet failed to ever be reproduced \cite{haibe2020importance}.

\paragraph{For some fields, the experiment is actually run by code}

Genetic epidemiology is a field within biology that
tries to determine the spread of heritable traits 
and their underlying biological mechanism.
For example, we know that lactose intolerance in adults is
caused by a decline in the production of lactose-degrading enzymes,
and is most commonly found in south-east asia and south africa \cite{storhaug2017country}.
The trait is caused by the genetic make-up, or 'genotype'
(see Table \ref{tab:definitions} for the definitions of terms used in this paper), 
of a person.
The trait, also called 'phenotype' (again, see Table \ref{tab:definitions}), 
in this example is lactose intolerance at adult age,
yet any human property, such as weight or height can be studied.
When an association between genotype and phenotype is found,
these associations are subsequently used to 
create so-called 'gene panels', 
where the gene causing 
an association is measured specifically, to, among others,
detect people at risk for the associated phenotype.
The rest of this section describes a genetic epidemiology study 
in more detail, with special focus on the computational experiment.

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{Karesuando_church.jpg}
  \caption{
    Picture of Karesuando's church,
    the village where the Northern Swedish Population
    Health Study started.
    From \cite{hopfner2005}
  }
  \label{fig:karesuando_church}
\end{figure}
\fi

The example study followed is a pseudorandomly selected paper
from \cite{ahsan2017relative}. The primary data used by that paper is
from a population study called the Northern Swedish Population
Health Study (NSPHS) that started in 2010 \cite{igl2010northern}. 
The approximately 1000 participants were initially mostly surveyed
about lifestyle \cite{igl2010northern} and follow-up studies
provided the type of data relevant for this paper, 
which are (1) the genotypes \cite{johansson2013identification},
(2) the phenotypes, in this case, concentrations of certain proteins in the 
blood \cite{enroth2014strong,enroth2015effect}.

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{1189px-Eukaryote_DNA-en.png}
  \caption{
    A cell has a nucleus that contains chromosomes. 
    Each of these chromosomes (46 in humans) consist out of DNA. 
    DNA itself conists out of 4 nucleotides, 
    as depicted by the horizontal sticks 
    with the colors red, yellow, green and blue.
    From \cite{sponk2012}
  }
  \label{fig:eukakyote_dna}
\end{figure}
\fi

The first type of primary data, the genotypes, 
consists out of single nucleotide polymorphisms (SNPs, pronounced 'snips').
A SNP has a name and a location on the DNA, at which there is a certain nucleotide.
DNA (organized into chromosomes and present in every (nucleated) cell, 
\iffalse
(see figure \ref{fig:eukakyote_dna}))
\fi
consists out of billions of nucleotides.
There are four types of nucleotides, 
called adenosine, cytosine, guanine and thyrosine, all commonly abreviated
as A, C, G and T respectively.
One SNP example is \verb|rs12133641|, which is a SNP located at position 
154,428,283, where 67 percent of the people within this study have an A,
and 33 percent have a G (also from \cite{ahsan2017relative}, Table S3).

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{DNA_to_protein.png}
  \caption{
    Parts of DNA (so-called 'genes') code for proteins. 
    The DNA, that always stays put in the cell's nucleus, 
    is transcripted to messenger RNA (mRNA).
    mRNA leaves the nucleus and its code gets translated to 
    a protein sequence.
    Near the start of a gene are regions that determine the amount
    of proteins produced (not shown in figure).
    Adapted from \cite{shafee2015}
  }
  \label{fig:dna_to_protein}
\end{figure}
\fi

The second type of data, the phenotypes, 
are concentrations of proteins in the blood. 
DNA contains the code for building proteins
\iffalse
(see Figure \ref{fig:dna_to_protein})
\fi
, as well as the rate
at which a protein is created. Some proteins end up in the blood and
their presence can be used to assess the health of an individual.
IL6RA is one such protein and its concentration may (and will, see below)
be associated with the SNP mentioned earlier.

The field of genetic epidemiology looks -among others- for
correlations between genetic data and biological traits.
For example, \cite{ahsan2017relative} shows that
SNP \verb|rs12133641| is highly correlated (p-value is $3.0^{-73}$
\iffalse
(see figure \ref{fig:ahsan2017relative_table_2_sub})
\fi
,
for $n$ = 961 individuals) with protein IL6RA.
The direction of the association is also concluded:
the more guanines (as opposited to adenines) are present at that SNPs location,
the higher concentration of IL6RA can be found in a human's blood
\iffalse
(see figure \ref{fig:ahsan2017relative_table_2_sub})
\fi
. The amount of variance that can be explained by an association (i.e.
the R squared value) is rarely 100 percent, which means that a trait (in
this case, the concentration of IL6RA) cannot be perfectly explained
from the genotype (in this case, SNP \verb|rs12133641|) alone. 
In this example, 43 percent of the variance 
can be attributed to an individuals' genotype
\iffalse
(see figure \ref{fig:ahsan2017relative_table_2_sub})
\fi
. Additional factors, 
such as the effect
of the environment (e.g. geographic location, time of day), 
lifestyle (e.g. smoking yes/no) or having a disease (e.g. diabetes) 
are needed to explain the additional variation.

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{ahsan2017relative_table_2_sub.png}
  \caption{
    An example result of a genetic epideological research.
    It shows that the SNP named rs12133641 (located at position 154,428,283
    of chromosome 1) is highly correlated (p value is 3.0 * 10e-73, 
    961 individuals) to the concentration of the protein IL6RA, as measured
    in blood. The table is a simplified result from \cite{ahsan2017relative}.
  }
  \label{fig:ahsan2017relative_table_2_sub}
\end{figure}
\fi

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{ahsan2017relative_s6.png}
  \caption{
    The relation between the genotype for SNP rs12133641 
    and the protein concentration of IL6RA is relatively strong.
    The X axis shows the the genotype of the individuals,
    where 0 denotes AA, 1.0 denotes AG and 2.0 denotes GG.
    The Y axis shows the concentration of the protein IL6RA 
    as found in the participants' blood. n = 1 denotes the number
    of SNPs that were determined to be involved.
  }
  \label{fig:ahsan2017relative_s6}
\end{figure}
\fi

The conclusions drawn from this paper,
may end up in the clinic.
For the sake of having a clear (yet fictitious) example,
let's assume that a high level of IL6RA 
is associated with a disease that develops later in life,
yet is preventable by lifestyle 
changes (see \cite{pope2003will} for an example in Alzheimer's disease).
Would this be the case, we can create a tailored 
experiment, called a gene panel, that specifically measures
SNP \verb|rs12133641|. 
If the gene panel shows an individual has two guanines, 
we know that this person is likelier to develop higher levels of
IL6RA and is likelier to benefit from the lifestyle changes.

From this simple example, it will be easier
to measure the level of IL6RA in the blood, than using a gene panel,
as blood tests are easier and cheaper.
However, there are associations published for many diseases,
in which one SNP (e.g. phenylketonuria) or many SNPs (e.g. \cite{bruce2009metabolic}) 
contribute to being more likely to develope a disease in the future. 
Here, the phenotype (having a disease in the future) 
is impossible to detect at the present
and associations found in earlier studies are used to create a gene panel.
As creating a gene panel is costly, those associations better be correct.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code is just as important as the paper}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\paragraph{The experiments within genetic epidemology works are done by code}

The experiment described above is run by code. 
It was code that detected the relationship between the genotype
(in this case, SNP \verb|rs12133641|) 
and the phenotype (in this case, the concentration of IL6RA).
To be more precise, it was code that read the data,
subsetted the data, removed outliers, 
performed the statistics and generated the plots.
All these steps are described in the paper (\cite{ahsan2017relative}), 
but the code itself is unpublished
and the paper does not mention any way to obtain it.
One could send an email to the correspending,
yet it is known that the reponse rate of correnspondence authors 
with a request for data in related fields is around 50\% 
\cite{manca2018non, teunis2015corresponding} 
(for the field of emergency medicine, these rates are 73\% \cite{o2003email}).

\paragraph{The goal of useful paradata is reproducible science}

The reason behind paradata is, among others,
to assure reproducible research \cite{huvila2022improving}.
Reproducibility (i.e. to reproduce the same results) 
and replicability (i.e. to reconclude a conclusion)
are fundamental characteristics of scientific studies \cite{patil2019visual}.
For computational science, theoretically, it should be relatively easy to 
reproduce an experiment, as all it takes is a computer, electricity,
an optional internet connection, the code and the data.
In practice, however, it appears that 
the academic culture to reproduce results 
has been lost over time in general \cite{peng2011reproducible},
with labs that embrace reproducibility (for example, \cite{barba2016hard})
are the exception.
To counter the lack of published code, 
it has been suggested to make the reproduction of research 
a minimal requirement for publication \cite{peng2011reproducible}.

\paragraph{Sensitive data}

A genetic epidemiologist works with sensitive data as well:
the genetic sequences of participants are private (\cite{clayton2019law}).
For research to be reproducible one needs both the code and the data
to reproduce the (hopefully) same results.
This problem is discussed in section \ref{sec:sensitive-data}.

\paragraph{Code holds the ground truth}

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{peng2011reproducible_fig_1.png}
  \caption{
    Levels of reproducibility, from \cite{peng2011reproducible}
  }
  \label{fig:peng2011reproducible}
\end{figure}
\fi

From a broader perspective, code should be published, as it holds the
ground truth of an experiment; it does the actual work.
The more complex the computation pipeline is, the easier it is
to have a mismatch between the article (that describes what the
code does) and the code (that actually does the work).
It is easy (tempting?) to overlook how easy it is to have a mismatch
between paper and code.

To illustrate how easy it is to get a mismatch between a paper
and the code, 
consider this example of a possible text in the fictional 
genetic epidemiology paper:

\begin{verbatim}
We compared the in-blood concentrations of IL6RA 
between subjects having AA versus subjects having GG at SNP rs12133641,
using an unpaired one-tailed T-test,
as we expect the average concentration of IL6RA for those with AA 
to be less than the average concentration in those with GG.
\end{verbatim}

We ignore the choices of words and style of this sentence: what is
important is the content: an unpaired one-tailed T test is performed.
Taking a look at the (in this case, the programming language R) code, 
we find the following line:

\begin{verbatim}
t_test_result <- t.test(subjects_with_aa, subjects_with_gg)
\end{verbatim}

The code is correct:
\verb|t.test| is indeed the name of an R function to do an unpaired T-test
and it is reasonable to assume that this code, when glancing over the code
of an article, is correct.

Here, however, we see a mismatch between the English description and the code:
by default, the R function \verb|t.test| does 
a \textbf{two}-tailed unpaired T-test.
A consequence of this fictional example is that the published p-values are
higher than needed, resulting in less significant findings, which
results in (needlessly) less conclusions drawn in this fictional paper.

Note that there is another side of this same coin:
correctly reporting the T-test in a reproducable ways
is known to be a problem, with only 80 out of a selected 179 papers
fully reporting the parity of T-test (i.e. paired or unpaired) 
\cite{weissgerber2018we}.
In those unresolved 99 papers, 
the code (if published) would clarify this.

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{baggerly2009deriving_fig_1a.png}
  \caption{
    Gene expression result due to an off-by-one error.
    The cells in the main table show how much mRNA is produced per cell 
    line (the columns) for different proteins (the rows).
    The colors above the cells, with the dendrogram, shows
    red for cells that do not respond (red) and do respond (blue) to an 
    antibiotic.
    The purple rectangles show samples that are duplicated.
    Note that, due to this, some
    samples that behave identically are both non-responders and responders.
    From \cite{baggerly2009deriving}
  }
  \label{fig:baggerly2009deriving}
\end{figure}
\fi

It is the paper that accompanies the code,
as it is the code that generates the results.
When humans are fallible and code gets bigger, the likelihood of
a mismatch between the English paper and the code increases.
We know that the most common errors are 
simple (and may legitimately fear that those simple errors are common) 
\cite{baggerly2009deriving},
and this example is not too far off from the
examples given in that paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Making code useful paradata}\label{sec:making-code-useful-paradata}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Again, useful paradata should be comprehensive, 
documented appropriately, the documentation should co-evolve with the data,
and the data should be friendly to computers \cite{huvila2022improving}.
In this section, these ideal properties of paradata are applied to code.
Code as paradata, however, has multiple uses as well,
which are to reproduce, to replicate and to 
extend a computational experiment \cite{benureau2018re}.
In this section, the recommendations how to make paradata useful
is applied to code and compared to existing best practices in software 
development, for the intended use of the code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Code must be usefully documented}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For code to be ideal paradata, it must be usefully documented \cite{huvila2022improving}.
What is 'useful' is determined by its intended use: 
to reproduce, replicate or extend \ref{sec:making-code-useful-paradata}.
The amount of documentation is, however, also determined
by the recommended practices for writing code in general.

For purposes of reproducing (same code, same data) or replicating (same code,
different data) a study, the minimal amount of documentation
is determined by software development best practices.
When an independent researcher reproduces or replicates an experiment, 
it is possible (even ideal!) that no code is read at all.
Due to this, one could argue that no documentation is needed.
However, all code should be documented 'adequately' \cite{peng2006reproducible},
ideally writing code in such a way that it becomes 
self-explanatory \cite{wilson2014best}
and for the remaining code to document the reasons behind it, 
its design and its purpose \cite{wilson2014best}.

For purposes of extending a study and its code, 
documentation becomes even more important,
as the code will be read and modified.
The extent of investing time in documenting code is recommended
to be proportional to the intended reuse \cite{pianosi2020successfully}
and there exists a clear relationship between the reuse 
of code and its documentation 
effort \cite{cosentino2017systematic,hata2015characteristics}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Code and documentation must align}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For code to be ideal paradata, 
its creation and documentation need to match, 
'not least because they shape each other' \cite{huvila2022improving}.
This quote resonates strongly with the idea of 
literate programming \cite{knuth1984literate},
in which documentation and code is developed hand-in-hand.
rOpenSci, a community that, among others, reviews programming
code (\cite{ram2013ropensci,ram2018community}), is 
one example of where literate programming is enforced:
code must have extensive runnable examples
as part of the documentation \cite{ropensci_2021_6619350}.
Developing code and documentation at the same time 
leads to both better code and documentation \cite{reenskaug1989environment}.
In general, when developing software, 
it is recommended to 
to write documentation while writing software,
as well as to include many examples \cite{lee2018ten}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Code must be extensive}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For code to be ideal paradata, it must be extensive.
As code has many properties, there are many recommendations on this aspect.

\subsubsection{... to reproduce an experiment}

\paragraph{Code should be hosted}

For code to be extensive enough to reproduce an existing experiment,
it should be distributed in standard ways \cite{peng2006reproducible}.
The standard way to publish code is to use a code hosting website.
There are multiple code hosting websites, 
with GitHub being the most popular one \cite{cosentino2017systematic}.
GitHub is a website where users can create dedicated pages (called 'repositories')
to upload code and interact with it.
The use of such code hosting websites
has increased strongly \cite{russell2018large},
accommodates collaboration \cite{perez2016ten}
and improves transparency \cite{gorgolewski2016practical}.
See \cite{cosentino2017systematic} for an extensive overview of
research conducted on GitHub.

\iffalse
An example from the author is the website \cite{bbbqarticleissue157},
that hosts part of the code for the paper \cite{bilderbeek2022transmembrane}, 
as shown in Figure \ref{fig:bbbqarticleissue157}.
\fi

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{bbbq_article_issue_157.png}
  \caption{
    A typical GitHub repository \cite{bbbqarticleissue157}, 
    hosting the code for a
    computational experiment, 
    as published with \cite{bilderbeek2022transmembrane}.
    Note that this repository has 72 commits, where a commit is a change
    to the code.
  }
  \label{fig:bbbqarticleissue157}
\end{figure}
\fi

\subsubsection{... to replicate an experiment}

\paragraph{Code must have extensive error handling}

Code must be more extensive when it is (intended to be) 
used on different data,
as then 'code must act as a teacher for future developers' \cite{sadowski2018modern}.
Error handling is one of the mechanisms to do so.
A simple example is when one (e.g. a BSc student) 
needs to calculate the variance of a distribution.
Commonly one needs to call a function to do so, 
e.g. \verb|var| in R, that needs the values of the distribution
as input.
However, to calculate the variance of a distribution, at least two values
are required (else a division by zero occurs in the calcation of the variance). 
When the variance of (zero or) only one value is requested,
an error message can helpfully indicate that the function 
needs at least two values to work as expected.

\subsubsection{... to extend an experiment}

In software development, 
changes are common and should be embraced \cite{beck2000extreme}.
When a scientist intends to reuse code by extending its functionality,
first the code will be assessed,
after which the code may be actually extended.

\subsubsubsection{Assessment of code}

\subsubsubsection{Extending code}

Code is just 'another kind of data' and should be designed 
as such, for example, by using tools to work on it \cite{wilson2022twelve}.

\paragraph{Using a tool for coding style}

A first example is to use a tool to enforce a coding style 
(e.g. the Tidyverse style guide \cite{wickham2019advanced} for R,
or PEP 8 \cite{van2001pep} for Python).
Following a consistent coding style improves software quality \cite{fang2001}.

\paragraph{Using a tool for complexity}

A second example is to use a tool to enforce a low cyclomatic complexity.
The cyclomatic complexity is approximately defined 
as the number of independent paths that
the code can be executed. 
For example, code that has only one \verb|if| statement
has a cyclomatic complexity of 2, as the condition within the \verb|if|
statement can be true or false,
resulting in the body of the \verb|if| statement being either
exectuted or ignored.
The cyclomatic complexity correlates with code complexity,
where more complex code is likelier to contain or give rise to bugs 
\cite{abd2018calculating,chen2019empirical,zimmermann2008predicting}

\paragraph{Code must be tested}

Coding errors are extremely common 
and contribute to the reproducibility crisis in science \cite{vable2021code}.

%.
%. How to make more published research true
%. NIH plans to enhance reproducibility
%. The replication crisis in epidemiology: snowball, snow job, or winter solstice?
%. The harm done to reproducibility by the culture of null hypothesis significance testing

\paragraph{Continuous integration helps assess code quality}

Some code hosts, among others, GitHub, 
allow the researcher/user to start software scripts upon uploading a change.
This is useful to determine the reproducibility of the code: 
one such script should run the code (on trivial data) 
to verify if the code works at all.
If the functionality of the code is lost, the researcher
can immediatly see which small change was the cause of this.
A formalized version of this workflow is called continuous integration (CI).
Using a CI service (such as GitHub Actions for GitHub), 
is known to significantly 
increase the number of bugs exposed \cite{vasilescu2015} and increases
the speed at which new features are added \cite{vasilescu2015}.
Using CI opens up the possibility to informally annotate features of
a repository. 
\iffalse
For example, Figure \ref{fig:build_badge} shows a 
build badge that indicates that a \LaTeX document (i.e. this paper)
could be built.
\fi

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{build_badge.png}
  \caption{
    A build badge for the GitHub repository of this article.
    The description indicates which process it is that the badge
    signals success (as in this case) or failure of. 
  }
  \label{fig:build_badge}
\end{figure}
\fi

\paragraph{Code coverage helps assess code quality}

Testing, in general, is an important mechanism to ensure
the correctness of code (an interesting example is \cite{rahman2020exploratory}
showing bugs in scientific software on the COVID-19 pandemic:
it recommends more testing).
The percentage of (lines of) code tested is called the code coverage.
Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
The non-profit organisation rOpenSci \cite{ram2018community},
which peer-reviews code,
has made it a prerequisite to have a code coverage of 100\%.
Also here, CI opens up the possibility to informally annotate the
code coverage. 
For example, build badges
\iffalse
(e.g. as shown in Figure \ref{fig:badge_codecov})
\fi
are visual indications on a website that can, among others,
show a certain R package's code has a 100\% code coverage.

Hence, code coverage is important metadata, that can automatically
be acquired when using code hosting website.

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[]{badge_codecov.png}
  \caption{
    A build badge for the code coverage of an R 
    package (\url{https://github.com/ropensci/beautier}).
    The umbrella signals that the code coverage is hosted by
    CodeCov (\url{https://codecov.io/}). 
    The text indicates which process it is that the badge
    signals, which in this case, is that the code is fully covered
    by tests.
  }
  \label{fig:badge_codecov}
\end{figure}
\fi

\paragraph{Adding a license}

Code hosting websites such as GitHub help developers to follow
best practices. One of these is to add a software license \cite{jimenez2017four},
so that it is clear that the software can be reused.

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{issue_ldpred2.png}
  \caption{
    An overview of the issues for the code for
    LDpred2 \cite{prive2020ldpred2}. Of the 310 issues, 300 have been closed.
    The first four issues have been replied to, as indicated by the
    talk balloon. The fifth issues has a label with the text 'feature request'
    to signal the issue type.
  }
  \label{fig:issue_ldpred2}
\end{figure}
\fi

\paragraph{Test-driven development}

TDD improves code quality \cite{alkaoud2018quality,janzen2006test}
and writing comment and documentation mixes well \cite{shmerlin2015document}.

\paragraph{Code must be reviewed to ensure the correct amount of documentation}

More than half of all scientists have their code 
rarely or never reviewed \cite{vable2021code},
although code reviews are known to
accelerate learning of the developers,
improve the quality of the code
and resulting to an experiment that is likelier to be reproducible
\cite{vable2021code}.
Within software development best-practices, 
code reviews are recommended as well \cite{wilson2014best}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Code must be computer friendly}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

There are multiple ways of publishing code,
that require increasingly more knowledge to do and result
in increasingly more reproducible experiments:
(1) publishing the code as text only, e.g. by printing
it in a Supplementary Materials, (2) publishing
the code on a hosted website, e.g on GitHub, with
added features (3) publishing 
the installed code within a virtual environment,
for example, using a Singularity container.
In this chapter, details of each of these ways will be described,
including advantages, disadvantages, as well as how to
make the code more FAIR data \cite{wilkinson2016fair}.

\paragraph{Publishing a virtual environment}

The most reproducible way of submitting the code of an experiment,
is by providing the code with all the (software) dependencies 
it needs installed in a virtual container.
This is close to the golden standard suggested by 
\cite{peng2011reproducible} 
\iffalse
(see also Figure \ref{fig:peng2011reproducible})
\fi
,
however, that article was written before virtualization existed.
Virtualization allows one to provide a standard environment,
that can be used on any other computer.
The best known examples of container software are Docker and Singularity.
Due to access rights, Singularity is the only viable option to run
an experiment on a 
high performance computing (HPC) environment (e.g. a computer cluster).

\paragraph{Containers allow for runnable code, for longer periods of time}

Unlike English, code is fragile in standing the test of time.
Containers can alleviate this, as these can be created at a point
in time and then run later. Because the container is stand-alone,
all software that the code depends on remain in their current/old state.
As with hosted code, an experiment may not run with newer versions
of libraries anymore, but a containerized experiment can run regardless, 
if one refrains from updating (or only does so when there are there problems with fixing).

Ideally, to re-do an experiment, one would need a one-liner to a container,
for example, \verb|singularity run my_experiment.sif run|.

\paragraph{Runnable code can precisely reproduce an experiment}

Containers allow a computation experiment to be highly reproducible:
given the same data, an experiment put into a container will give
the same results on different platforms, at least in theory.
In practice, differences may be observed when peripheral factors
are different, such as the random numbers as generated by an operating
system, or data that are downloaded from additional online sources.

\paragraph{Runnable code allows others to re-use an experiment}

Containers allow a computation experiment to be highly re-usable,
as any scientist can work with it and tweak code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preserving code}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Preserving code by uploading it to a code hosting webite
has many advantages.

\paragraph{Hosted code has important metadata}

Hosted code has many additional features, such as
a version-controlled history, continuous integration scripts
and automatically generated metrics.
A code hosting website produces and shares these metadata.
These metadata can help to determine the
reproducibility, correctness and quality of code
of computational experiments.
Below, those features and why these are important are discussed.

\paragraph{Hosted code keeps a history of changes}

Hosted code commonly has a commit history. For example, GitHub
uses a version control system called \verb|git| (hence its name).
In practice, this means that when a change is made to the code,
a new version is created. In the case that the change was harmful,
one can go back to an earlier version and continue again from there.
It is a general recommendation to put version control
on all human-produced data \cite{wilson2014best},
as well as openly working on the code from the start \cite{jimenez2017four}.

\paragraph{A commit history helps assess honest reporting}

Hosted code improves transparency \cite{gorgolewski2016practical},
as, among others, the history of changes (i.e. the git commit history), 
helps to assess the honesty in the reporting of the code. 
Ideally, the full version history of the code 
has been generated, by hosting the code directly at its inception,
so that the ontology of the experiment can be viewed.
However, in some repositories, only the final version of the code has been
posted, which means all historical information is lost.
The code history has many important properties,
as, for example, 
it can be observed how much statistical tests have really/actually been done 
and see if all are reported:
failing to reporting a (usually non-significant) statistical tests is 
one of the many degrees of freedom 
in p-value hacking \cite{wicherts2016degrees}.

\paragraph{Metrics help assess the honest division of labour}

Most code hosts, among others, GitHub and GitLab,
automatically keep track of certain metrics.
One such metric is the amount of code each contributor has made.
In that way, authors involved in the writing of the software
are honestly acknowledged. 
\iffalse
Figure \ref{fig:daisie_contributors}
shows the number of commits the authors of DAISIE \cite{etienne2020daisie}
have made to its code.
\fi
The distribution of contributed code per author 
is useful metadata, that can automatically
be acquired when using code hosting website.

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{daisie_contributors.png}
  \caption{
    An overview of the commits that multiple contributors have
    made, in this case, to the code of DAISIE \cite{etienne2020daisie}
  }
  \label{fig:daisie_contributors}
\end{figure}
\fi

\paragraph{Suggestions for knowledge management}

Here are multiple suggestions to make containers with runnable code FAIR.
To make such containers findable, the most common 
environments (Syslabs and DockerHub) to upload containers must be indexed.
Uploading a container to those environments is, however, limited by
a restricted amount of free storage, with the possible consequence that
those containers are uploaded somewhere where they escape proper indexing.
Having containers uploaded at the existing websites ensures they are
accessible, interoperable and reusable.
One could argue however, that the scientific community need its own,
similar website to upload containers, as the metadata for these
sites are limited to free-text comments and tags (DockerHub)
or not metadata at all (Syslabs).

Code is harder to preserve than an English text
and preserving code is rarely done \cite{barnes2010publish}.
Although code is the primary actor in computational experiments,
there is no incentive to submit code alongside a publication.
Most academic journal do not require authors to submit their code,
nor is the submitted code peer reviewed, 
rOpenSci \cite{ram2018community} being the pleasant exception.

Code used in academic research complies badly with FAIR principles,
where unpublished code is the worst offender.
However, any idealistic researcher does not have the
tools to following FAIR principles in an exemplary way:
there are no standards in how a scholarly article
refers to academic code (i.e. in a machine-friendly way).
Creating a (Singularity) container is the best a researcher
can do to ensure his/her code is reusable,
yet, there are no incentives to make code accessible, interoperable
or re-usable, with standarized metadata attributes being absent.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sensitive data}\label{sec:sensitive-data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Next to the code, it is the data used in an experiment 
that must be made available for an experiment 
to be called 'reproducible' \cite{peng2006reproducible}.
When the input data are sensitive, however, and thus cannot be shared with
an independent researcher, one thus cannot reproduce the
experiment.

The solution for this problem is simple:
supply a simulated (also called 'analytical' \cite{peng2006reproducible}) dataset
together with the code.
In the case of genetic epidemiology this would mean
simulated genotypes and associated phenotypes,
as can be done with the \verb|plinkr| R package \cite{plinkr}.
This solution, however, is known to all developer that follow
good practices when writing code, as public (i.e. simulated) 
datasets are needed for automatic testing of code.
One extra benefit of simulated data is that these can be used
as a benchmark, as slightly different analyses should give 
similar conclusions.

For research to be reproducible, one ideally has access to
both the data used and the code.
In some fields, such as genetic epidemiology, the data are
sensitive, hence cannot be released,
yet there are methods being devised to run code on sensitive
data with assured privacy \cite{zhang2016review,azencott2018machine}.
Additionally, it follows software development's best practices 
(especially the practice to test code automatically) 
to release a simulated/public dataset, with
the additional benefit that this dataset can be used for comparisons.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In a perfect world, all code has the characteristis of ideal paradata
and is written from software development best practices.
This section discusses the problems that arise by doing so.

\paragraph{Invest in training}

One needs to be trained to know the best practices,
before these can be applied. 
Articles that suggest these best practices, however, claim that this initial
investment pays off.
Code reviews are a good way to accelerate the 
learning of team members \cite{vable2021code}.

\paragraph{Code needs maintanance}

Code needs maintanance:
the libraries that are used by an experiment
will change over time, inevitably causing the experiment to fail
over time. 
The degradation of software is a known feature for nearly 
four decades (see the definition of 'bit rot' at \cite{steele1983hacker}),
here a more modern (and in this context superior)
definition is used: software collapse \cite{hinsen2019dealing},
in which software fails due to dependencies on other 
software.
Where that paper recommends using older software foundations 
with multiple implementations \cite{hinsen2019dealing},
this paper recommends a modern solution that potentially has more 
longevity:
to publish the code installed in a virtual environment
(including the dependencies used at that time), as described below.
Write code that will stand the test of time perfectly
is deemed 'impossible' \cite{benureau2018re}.

\paragraph{Authors can be contacted}

There is another problem that authors of academic code face,
which is that they may be contacted with a user 
asking for technical report 
(against which a 'no' is just fine \cite{barnes2010publish})
or -more relevant to the research- sending in a bug report.
One solution for the author is to ignore such emails (as is done in a third
of all cases \cite{teunis2015corresponding}):
it can be argued that no energy should be wasted on published code
and work on something new instead 
(see \cite{barnes2010publish} for a better way to deal with the problem).

\paragraph{How to handle bugs?}

However, upon the find of a new bug, the question is:
does all the research that use that software still 
result in the same conclusions?
To prevent authors chosing to hide bug reports, 
a code hosting website usually has a public place to report bugs,
to increase transparency in the handling of bug reports.

\paragraph{The problems with runnable code}

There are some problems with providing a container with a runnable container.
First, a container is several gigabytes and need to be stored somewhere
online. 
Ideally, for Singularity, this would be at the main container hosting
site called Syslabs. Using Syslabs is free upon registration, 
yet there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
be entered, such as the URL of the article, nor the URL of hosted code.
Also, journals do not request hosted code, nor runnable code.

% \paragraph{The drawbacks of publishing code}

Publishing code may be disadvantageous for an author.
For science, code should be published, 
this allows reproducible research 
(again, see \cite{haibe2020importance} for a tragic example).
For an author, publishing code alongside an experiment opens up
the possibily to receive questions regarding that code.
Note, however, that not publishing code may put 
oneself in the focus of attention
and -after much effect by others reproducing an incorrect result-
at the cost of a scientific career \cite{baggerly2009deriving}.

% \paragraph{The drawbacks of publishing version-controlled code}

Is it worth it to publish version-controlled code?
For an author, 
there is additional training involved, and also here,
publishing code alongside an experiment opens up
the possibily to receive questions regarding that code,
as well as other interactions, including helpful contributions.

% \paragraph{The drawbacks of publishing a running version of code}

Is it worth it to archive running versions of code?
For an author, 
there is additional training involved.
There is no FAIR infrastructure for Singularity containers.

% \paragraph{Still need the same data to do a reproducible experiment}

% \paragraph{Recommendations}

Being able to re-do an experiment is a core principle of the scientific method.
Publishing only a paper about a computational experiment is not enough,
as the results are too likely to mismatch that English description.
Journal should make it mandatory for authors
to publish their code alongside a computational experiment.
Authors should follow the FAIR principes for their code as well,
as can be done using the infrastructure as supplied by, 
for example, GitHub and GitLab, opening up new angles in
metascience regarding computational experiments.
Knowledge managers should create the infrastructure for the preservation
of runnable experiments, to allow scientist to upload a Singularity
container, so these are as well following the FAIR principles.

\iffalse
\rowcolors{2}{gray!10}{gray!40}
\begin{table}[h]
  \begin{tabular}{|p{2cm}|l|}
    \hline
    \textbf{Type of code} & \textbf{Recommendation} \\
    \hline
    No code       & Disallow papers that do not supply code publicly \\
    \hline
    Raw code      & Don't publish raw code as text, host it on a code hosting website \\
    \hline
    Hosted code   & Use a standarized badge for build status \\
                  & Use a standarized badge for code coverage \\
                  & Standardize to link between paper and code host \\
    \hline
    Runnable code & Allow to upload containers without limits \\
                  & Website to host containers must allow for annotation \\
                  & Use a standarized badge for the container to be functioning  \\
                  & Standarize to link between container, paper and code repository \\
    \hline
  \end{tabular}
  \caption{Recommendations described in this paper}
  \label{tab:recommendations}
\end{table}
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This paper started with some suggestions to 
make paradata useful for data re(use):
paradata should be comprehensive, comprehensively documented
and friendly to computers \cite{huvila2022improving}.
When applying these general recomendations to code, 
this list can be phrased more precisely:
code should be comprehensive in supplying 
automatically generated metadata (such as commit history and code coverage).
The documentation should be as extensive as recommended by the 
software development literature. To be computer friendly,
code should be able to run on at least one independent entity,
such as a GitHub Actions server or from a Singularity container.

\paragraph{Final goal: runnable code}

When research truly needs to be reproducible, putting the code 
of an experiment into a container is a good solution, as it 
can stand the test of time better.
Creating such a container, however, requires more skill
that -as of today- is not rewarded,
partly as containers follow FAIR data principles only poorly.
Runnable code is the pinnacle of reproducible research.

\paragraph{First step: publish code alongside papers}

% \paragraph{Final word}

The world of science would be a more open, humble, trustworthy, truthful
and helpful would the code that accompanies a scientific paper
be treated like a first class citizen. 
Doing so, however, is yet to be rewarded
and still both of the two scientists at the start of this paper 
can provide a good rationale for their behavior.
This will change when reward incentives are put into place 
that reward making paradata useful.
For code specifically, in any computational field,
the rewards are even higher, as reproducibility should again be 
a cornerstone in science.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Accessibility}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This article and its metadata can be found at 
\begin{sloppypar}\url{https://github.com/richelbilderbeek/chapter_paradata}\end{sloppypar}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Vancouver style
\bibliographystyle{unsrtnat}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\section{Supplementary materials}

% Figures start from one and are prepended with an S
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

% Tables start from one and are prepended with an S
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{This paper}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This paper has been produced following all recommendations in it:
it's \LaTeX code is hosted on GitHub 
at \url{https://github.com/richelbilderbeek/chapter_paradata}.
Due this, one can, among others, see the complete version history of this paper.
This alone could allow research to be done on, for example,
the effect of reviewers' feedback on a first manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Funding}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This chapter has not been supported by funding.

% Duran et al., grounds for trust
% [A] process is epistemically opaque relative to a cognitive agent X at time t
% just in case X does not know at t all of the epistemically relevant elements of
% the process (Humphreys 2009, 618)
% 
% On Computational reliability
% 
% (CR) if S’s believing p at t results from m, then S’s belief in p at t is justified.
% where S is a cognitive agent, p is any truth-valued proposition related to the
% results of a computer simulation, t is any given time, and m is a reliable com‑
% puter simulation.
%
%
% 4 Sources for Computational Reliabilism
%
% 1. Verification and validation methods
% 2. Robustness analysis for computer simulations
% 3. A history of (un)successful implementations
% 4. Expert knowledge